---
title: "Preliminary_analysis"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

To get Some Sense of Data and at the same build some groundwork for Data Collection

```{r }
library(dplyr)
library(plyr)
library(janitor)
Covid_Nursing_Home <- read.csv("~/Health_IT/Covid_Nursing_Home.csv", header=FALSE)
Covid_Nursing_Home <- Covid_Nursing_Home %>%
  row_to_names(row_number = 1)
Covid_Nursing_Home <- rename(Covid_Nursing_Home, c("Provider State"="State", "Residents Total COVID-19 Deaths"="total_deaths"))

Covid_Nursing_Home$total_deaths <- as.numeric(Covid_Nursing_Home$total_deaths)
Covid_Nursing_Home <- na.omit(Covid_Nursing_Home)

#mean number of deaths in nursing homes due to covid
mean(Covid_Nursing_Home$total_deaths)

```

## Mean Deaths in florida and distribution

You can also embed plots, for example:

```{r}
Fl <- Covid_Nursing_Home %>% filter(State=="FL")
library(ggplot2)
Fl$total_deaths <- as.numeric(Fl$total_deaths)
ggplot(Fl, aes(x = State, y = total_deaths)) +
  geom_boxplot()
Fl <- na.omit(Fl)
mean(Fl$total_deaths)

```

## Mean Number of Deaths at Nursing Homes in FLorida : 1.77

Box plot suggests to investigate nursing homes with higher deaths as clearly distribution is heavily left skewed 

As a benchmark let's investigate more than 20 deaths

```{r}
death_20 <- Covid_Nursing_Home %>% filter(total_deaths>20)
# number of rows where death is more than 20 : 5566 from the original 195,000 rows
#1/40th of the nursing homes have death greater than 20

ggplot(death_20, aes(x = State, y = total_deaths)) +
  geom_boxplot()

```


## It can  be inferred from the box plot that states like Florida, South Carolina and Minnesota needs to be properly investigated for IT capabilities and also rest of them

# ANOVA
```{r}
library(statsr)
res.aov <- aov(total_deaths ~ State, data = death_20)
# Summary of the analysis
summary(res.aov)

```

Assuming deaths are in normal distribution for each state, ANova suggests that at least two of the states have varying means from one another.

### Main Conclusions
Distribution is heavily left skewed meaning some if the nursing homes have more death than rest of them - will check IT there

Mean when more than 20 deaths is : 34

States like South Carolina, Florida, Minnesota are on investigation radar





